---
title: "Quickstart guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quickstart guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(convo)
```

## Define

Define stubs and optional metadata in YAML:

```{yaml, code=xfun::read_utf8(system.file("", "ex-convo.yml", package = "convo")), eval = FALSE}
```

Read into R:

```{r}
filepath <- system.file("", "ex-convo.yml", package = "convo")
convo <- read_convo(filepath)
print(convo)
```

## Evaluate & Compare 

Evaluate a set of names (e.g. variable or file names) against the `convo` object to find violations:

```{r}
col_names <- c("ID_A", "IND_A", "XYZ_D", "AMT_B", "AMT_Q", "ID_A_1234", "ID_A_12")
evaluate_convo(convo, col_names)
```

Compare stub lists and identify new potential entries (stubs used in variables but not in controlled vocabulary):

```{r}
convo_colnames <- parse_stubs(col_names)
convo_colnames
compare_convo(convo_colnames, convo, fx = "setdiff")
```

(Note that if your names are separated by a different delimiter than `_`, you may pass that to `parse_stubs()` and most other functions shown in this demo using the `sep =` argument.)

If desired, newly uncovered stubs can be manually added to the `convo`:

```{r}
convo2 <- add_convo_stub(convo, level = 2, stub = "B", desc = "Type B")
convo2 
```

Or, alternatively, the `compare_convo()` function can accept a `"union"` option to merge needed new stubs between two objects:

```{r}
convo_union <- compare_convo(convo_colnames, convo, fx = "union")
convo_union
```

## Record

After conducting set operations on a `convo` as shown above, a new `convo` YAML specification can be written back out to YAML:

```{r}
write_convo(convo_union, filename = "new-convo.yml", path = tempdir())
```

```{yaml, code= file.path(tempdir(), "ex-convo.yml"), eval = FALSE}
```

## Validate 

Generate a `pointblank` agent for data validation:

```{r}
filepath <- system.file("", "ex-convo.yml", package = "convo")
convo <- read_convo(filepath)
agent <- create_pb_agent(convo, data.frame(IND_A = 1, IND_B = 5, DT_B = as.Date("2020-01-01")))
pointblank::interrogate(agent)
```

Or create a `pointblank` YAML file for portability: 

```{r echo = FALSE}
tmp <- gsub("\\\\", "/", tempdir())
write_pb(convo, c("IND_A", "AMT_B"), filename = "convo-validation.yml", path = tmp)
```

```{r eval = FALSE}
filepath <- system.file("", "ex-convo.yml", package = "convo")
convo <- read_convo(filepath)
write_pb(convo, c("IND_A", "AMT_B"), filename = "convo-validation.yml", path = tempdir())
```

```{yaml, code=xfun::read_utf8(file.path(tempdir(), "convo-validation.yml")), eval = FALSE}
```

## Document

Make data dictionary:

```{r}
vars <- c("AMT_A_2019", "IND_C_2020")
desc_df <- describe_names(vars, convo, desc_str = "{level1} of {level2} in given year")
DT::datatable(desc_df)
```

Visualize data as controlled vocabulary components:

```{r}
vbls <-  c("AMT_A_2019", "AMT_B", "AMT_C", "IND_A", "IND_B_2020")
vbls_df <- parse_df(vbls)
viz_names(vbls_df)
```

Describe an overall `convo` specification:

```{r}
DT::datatable(describe_convo(convo))
```

Optionally include the "contracts" (or validation checks) in the documentation (also powered by `pointblank`):

```{r}
DT::datatable(describe_convo(convo, include_valid = TRUE))
```

## Beyond Column Names

`convo` can manage controlled vocabularies beyond column names, as well. For example, `convo` can help document files is a large project. 

```{r}
filenames <- c("analysis/validation-a.Rmd",
               "analysis/validation-b.Rmd",
               "analysis/analysis.Rmd",
               "analysis/report.Rmd",
               "src/script-a.sql",
               "src/script-b.sql")
filenames_clean <- gsub("\\.[A-Za-z]+$", "", basename(filenames))
parse_stubs(filenames_clean, sep = "-")
```




